---
title: "Zillow Housing Data Wrangling and Cleaning"
author: "Jomar Sastrillo"
date: "January 21, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rjson)
library(stringr)
```

## Exploring the housing data
We first explore the structure of the data provided.

```{r}
# The first item is a description of every field that we could possibly find in the data.
data_dictionary <- read_csv("../data/DataDictionary.csv", col_types=cols())

all_available_metrics <- fromJSON(file = "../data/all_available_metrics.json")

# Every field for every level (level=state, etc..)
fields_per_level <- fromJSON(file = "../data/fields_per_level.json")
```

We want to know which variables are available for which resolution of housing data.

```{r}
# A data frame containing whether the field is available for the given level
data_avail <- fields_per_level %>% 
    map_dfc(function(x){all_available_metrics %in% x}) %>% 
    mutate(field = all_available_metrics) %>% 
    select(field, everything())

data_avail %>% head()
```

Find the regional views which have all the fields.

```{r}
data_avail %>% 
    filter( City==T, County==T, Metro==T, Neighborhood==T, State==T, Zip==T )
```

## Extract the data

```{r}
state_homes <- read_csv(file = '../data/State_time_series.csv', col_types = cols())
zip_homes <- read_csv(file = '../data/Zip_time_series.csv', col_types = cols())
county_homes <- read_csv(file = '../data/County_time_series.csv', col_types = cols())
city_homes <- read_csv(file = '../data/City_time_series.csv.csv', col_types = cols())
```

There's a lot of variables here. First, we make sure that the correct number of states are selected.

```{r}
sel_variables <- state_homes %>% 
    filter(RegionName!='UnitedStates') %>% 
    #filter(Date >= '2000-01-01') %>% 
    summarize_all( funs( sum(is.na(.)) ) ) %>% 
    gather(variable, na_count, 1:79) %>% 
    arrange(na_count) %>% 
    slice(3:14) %>% 
    pull(variable)
```

```{r}
sel_variables
```


```{r}
final_homes <- state_homes %>% 
    filter(RegionName!='UnitedStates') %>% 
    filter(Date >= '2000-01-01') %>% 
    rename(state_name = RegionName) %>% 
    select( c('Date', 'state_name', sel_variables) )
```

We write the wrangled results to a file

```{r}
write_csv(final_homes, path = "../app/final_homes.csv", col_names = T)
```

```{r}
final_homes
```

```{r}
final_homes %>% 
    select(Date, state_name, qq=ZHVI_3bedroom) %>% 
    group_by(state_name) %>%
    arrange(desc(Date)) %>%
    slice(1) %>% 
    right_join(state_boundaries, by = 'state_name') %>% 
    filter_all( all_vars(!is.na(.)) )
```

```{r}
final_homes %>% 
    filter(state_name=="Maine") %>% 
```

```{r}
read.csv("../app/final_homes.csv", stringsAsFactors = FALSE) %>% 
    mutate(Date = date(Date), state_name=as.factor(state_name)) %>% 
```